---
title: "Chap5_The_tidyverse"
output: html_notebook
---

* * *

# 5.0 vector to data frame 

Up to now we have been manipulating vectors by **reordering** and **subsetting** them through **indexing**. However, once we start more advanced analyses, the _preferred unit for data storage is not the vector_ but the **data frame**. In this chapter we learn to work directly with **data frames**, which greatly facilitate the organization of information.. We will be using data frames for the majority of this book. We will focus on a specific data format referred to as tidy and on specific collection of packages that are particularly helpful for working with _tidy_ data referred to as the `tidyverse`.

We can load all the tidyverse packages at once by installing and loading the tidyverse package:

```{r}
install.packages("tidyverse")
library(tidyverse)
```

###### >> conlicts : name space가 겹친 경우 (chap4 참고, {package_name}::{function_name})

We will learn how to implement the tidyverse approach throughout the book, but before delving into the details, in this chapter we introduce some of the most widely used tidyverse functionality. starting with the `dplyr` package for **manipulating data frames** and the `purrr` package for **working with functions**. Note that the tidyverse also includes a **graphing package**, `ggplot2`, which we introduce later in Chapter 8 in the Data Visualization part of the book; the `readr` package discussed in Chapter 6; and many others. In this chapter, we first introduce the **concept of tidy data** and then demonstrate how we use the tidyverse to work with data frames in this format.

> major packages in `tidyverse`

1. `dplyr` : manipulating data frames
2. `purrr` : working with function (we already use `gsup` function)
3. `ggplot2` : graphing package (chap8)
4. `readr` : importing module 


* * *

# 5.1 Tidy data

We say that a data table is in tidy format if each **row** represents _one observation_ and **columns** represent the _different variables_ available for each of these observations. The `murders` dataset is an example of a tidy data frame.

``` 
tidy data 
1. each raw >> one observation
2. each column >> different variables 
```

```{r}
#>        state abb region population total
#> 1    Alabama  AL  South    4779736   135
#> 2     Alaska  AK   West     710231    19
#> 3    Arizona  AZ   West    6392017   232
#> 4   Arkansas  AR  South    2915918    93
#> 5 California  CA   West   37253956  1257
#> 6   Colorado  CO   West    5029196    65
```


Each row represent a state with each of the five columns providing a different variable related to these states: name, abbreviation, region, population, and total murders.

To see how the same information can be provided in different formats, consider the following example:

```{r}
#>        country year fertility
#> 1      Germany 1960      2.41
#> 2  South Korea 1960      6.16
#> 3      Germany 1961      2.44
#> 4  South Korea 1961      5.99
#> 5      Germany 1962      2.47
#> 6  South Korea 1962      5.79
#> 7      Germany 1963      2.49
#> 8  South Korea 1963      5.57
#> 9      Germany 1964      2.49
#> 10 South Korea 1964      5.36
#> 11     Germany 1965      2.48
#> 12 South Korea 1965      5.16
```

```
Q) Does each columns seperated by tap or space? maybe tap..? I'm not sure. 
```

This tidy dataset provides fertility rates for two countries across the years. This is a tidy dataset because each row presents one observation with the three variables being county, year and fertility rate. However, this dataset originally came in another format and was **reshaped** for the `dslabs` package. Originally, the data was in the following format:

```{r}
#>       country 1960 1961 1962 1963 1964 1965
#> 1     Germany 2.41 2.44 2.47 2.49 2.49 2.48
#> 2 South Korea 6.16 5.99 5.79 5.57 5.36 5.16
```

The same information is provided, but there are two important differences in the format: 

> 1) each row includes several observations and 
> 2) one of the variables, year, is stored in the header. 

For the tidyverse packages to be optimally used, data need to be **reshaped into tidy format**, which you will learn to do in the **Data Wrangling** part of the book. Until then, we will use example datasets that are already in tidy format.

```
Data Warngling(=Data munging) : Transforming and mapping from one "raw" data form into another format. 
Data Parsing(=syntax analysis) : Analysing a string of symbols conforming to the rules of a formal grammar and making the parse tree.
** parse = part (of speech)
```

Although not immediately obvious, as you go through the book you will start to appreciate the advantages of working in a framework in which functions use _tidy formats for both inputs and outputs_. You will see how _this permits the data analyst to focus on more important aspects_ of the analysis rather than the format of the data.

* * *

# 5.2 Exercise

1. Examine the built-in dataset co2. Which of the following is true:

A. co2 is tidy data: it has one year for each row.

B. co2 is not tidy: we need at least one column with a character vector.

C. co2 is not tidy: it is a matrix instead of a data frame.

 _> isn't it correct?_

**D. co2 is not tidy: to be tidy we would have to wrangle it to have three columns (year, month and value), then each co2 observation would have a row.**


```{r}
data("co2")
co2
```

> each row : time series data of each year  
> each column : each month 

```{r}
# 왜 head(co2)는 header를 출력해주지 않는가? 
head(co2)
```

```{r}
class(co2)
# 1. dataframe이 아니기때문 
# 2. matrix (ts가 matrix의 일종인듯)로 인식되기 때문 
```
> ts : [time-series object](https://stat.ethz.ch/R-manual/R-devel/library/stats/html/ts.html) 



2. Examine the built-in dataset ChickWeight. Which of the following is true:

A. ChickWeight is not tidy: each chick has more than one row.

**B. ChickWeight is tidy: each observation (a weight) is represented by one row. The chick from which this measurement came from is one the variables.**

C. ChickWeight is not a tidy: we are missing the year column.

D. ChickWeight is tidy: it is stored in a data frame.


```{r}
data("ChickWeight")
head(ChickWeight,10)
class(ChickWeight)
```



3. Examine the built-in dataset BOD. Which of the following is true:

A. BOD is not tidy: it only has six rows.

B. BOD is not tidy: the first column is just an index.

**C. BOD is tidy: each row is an observation with two values (time and demand)**

D. BOD is tidy: all small datasets are tidy by definition.


```{r}
data("BOD")
BOD
```



4. Which of the following built-in datasets is tidy (you can pick more than one):

A. BJsales

**B. EuStockMarkets**

**C. DNase**

**D. Formaldehyde**

**E. Orange**

F. UCBAdmissions


```{r}
data("BJsales",EuStockMarkets,DNase,Formaldehyde,Orange,UCBAdmissions)

head(BJsales)
head(EuStockMarkets)
head(DNase)
head(Formaldehyde)
head(Orange)
head(UCBAdmissions)
```

```{r}
#dataset_name = c(BJsales,EuStockMarkets,DNase,Formaldehyde,Orange,UCBAdmissions)
#for (i in dataset_name){
#  print(head(i,5))
#}
```


```
>> it's not working
1) dataset_name이라는 vector에 변수명을 string(""이용)으로 저장하는 경우 => 그저 string으로 for-loop을 돌게되어 무의미
2) string이 아니게 저장하는 경우 => 이상한 방식으로 for loop을 돎... 
>> 원인 : numeric이나 character가 아니고, data frame이나 matrix를 vector로 묶으려고 하면 list가 되어버림. 
 - 변수형에 따라 for-loop 실험해보기 
```


***

# 5.3 Manipulating data frames 

The `dplyr` package from the tidyverse introduces functions that perform some of the most common operations when working with data frames and uses names for these functions that are relatively easy to remember. For instance, to change the data table by adding a new column, we use `mutate`. To filter the data table to a subset of rows, we use `filter`. Finally, to subset the data by selecting specific columns, we use `select`.

1) `mutate` : change the data table by adding a new column
2) `filter` : filter the data table to a subset of rows 
3) `select` : subset the data by selecting specific columns 

## 5.3.1 Adding a column with `mutate`

We want all the necessary information for our analysis to be included in the data table. So the first task is to add the murder rates to our murders data frame. The function `mutate` takes the data frame as a first argument and the name and values of the variable as a second argument using the convention name = values. So, to add murder rates, we use:


```{r}
library(dslabs) #Data Science Labs 
data("murders")
head(murders)
```
```{r}
murders <- mutate(murders, rate = total / population * 100000)
head(murders)
# successfully insert the column 
```

Notice that here we used `total` and `population` inside the function, which are objects that _are not defined in our workspace._ But why don’t we get an error?

This is one of `dplyr`’s main features. Functions in this package, such as `mutate`, know to look for variables in the data frame provided in the **first argument**. In the call to `mutate` above, total will have the values in murders$total. This approach makes the code much more **readable**.

```
* dplyr는 data frame의 first argument를 일종의 변수명으로 인식한다. 
```

We can see that the new column is added:
```{r}
head(murders)
#>        state abb region population total rate
#> 1    Alabama  AL  South    4779736   135 2.82
#> 2     Alaska  AK   West     710231    19 2.68
#> 3    Arizona  AZ   West    6392017   232 3.63
#> 4   Arkansas  AR  South    2915918    93 3.19
#> 5 California  CA   West   37253956  1257 3.37
#> 6   Colorado  CO   West    5029196    65 1.29
```

Although we have overwritten the original `murders` object, this does not change the object that loaded with `data(murders)`. If we load the `murders` data again, the original will overwrite our mutated version.



## 5.3.2 Subsetting with `filter`

Now suppose that we want to filter the data table to only show the entries for which the murder rate is lower than 0.71. To do this we use the `filter` function, which takes the **data table as the first argument** and then the **conditional statement as the second**. Like mutate, we can use the unquoted variable names from murders inside the function and it will know we mean the columns and not objects in the workspace.

> filter(data table, conditional statement)
> conditional statement : just like 'filter' function of excel. We should type the exact name of variable. 


```{r}
filter(murders, rate <= 0.71)
```



## 5.3.3 Selecting columns with `select`

Although our data table only has six columns, some data tables include hundreds. If we want to view just a few, we can use the **dplyr** `select` function. In the code below we select three columns, assign this to a new object and then filter the new object:

> purpose of `select` : select a few columns from full data frame 


```{r}
new_table <- select(murders, state, region, rate)
filter(new_table, rate<=0.71)
```

In the call to `select`, the first argument `murders` is an object, but state, region, and rate are variable names.

> select(data table, name of column #1, name of column #2, ...)


***

# 5.4 Exercises

1. Load the **dyplr** package and the muders dataset.


```{r}
library(dplyr)
library(dslabs)
data(murders)
```

You can add columns using the **dyplr** function `mutate`. This function is aware of the column names and inside the function you can call them unquoted:


```{r}
murders <- mutate(murders, population_in_millions = population / 10^6)
head(murders)
```


 We can write `population` rather than `murders$population`. The function `mutate` knows we are grabbing columns from `murders`.

 Use the function `mutate` to add a murders column named `rate` with the per 100,000 murder rate as in the example code above. Make sure you redefine `murders` as done in the example code above(murders < - [your code]) so we can keep using this variable. 
 


```{r}
murders <- mutate(murders, rate = total / population * 10^5)
head(murders)
```



2. If `rank(x)` gives you the ranks of `x` from **lowest to highest**, `rank(-x)` gives you the ranks from **highest to lowest**. Use the function mutate to add a column rank containing the rank, from highest to lowest murder rate. Make sure you redefine `murders` so we can keep using this variable.

```{r}
murders <- mutate(murders, rank = rank(-rate))
head(murders)
```

```
rank(column name) = ranks from lowest to highest
rank(-column name) = ranks from highest to lowest 
```



3. With **dplyr**, we can use `select` to show only certain columns. For example, with this code we would only show the states and population sizes:

```{r}
select(murders, state, population) %>% head()
```

```
%>% : pipe (just like in terminal)
```

Use `select` to show the state names and abbreviations in murders. Do not redefine murders, just show the results.

```{r}
select(murders, state, abb) %>% head()
```



4. The **dplyr** function `filter` is used to choose specific rows of the data frame to keep. Unlike `select` which is for columns, `filter` is for rows. For example, you can show just the New York row like this:


```{r}
filter(murders, state == "New York")
```

```
select : selecting the columns
filter : selecting the rows 
```

You can use other logical vectors to filter rows.

Use `filter` to show the top 5 states with the highest murder rates. After we add murder rate and rank, do not change the murders dataset, just show the result. Remember that you can filter based on the rank column.

```{r}
filter(murders, rank<= 5 )
```

5. We can remove rows using the `!=` operator. For example, to remove Florida, we would do this:

```{r}
no_florida <- filter(murders, state != "Florida")
```

Create a new data frame called `no_south` that removes states from the South region. How many states are in this category? You can use the function `nrow` for this.

```{r}
nrow(murders)
```

```{r}
no_south <- filter(murders, region != "South") 
nrow(no_south)
```



6. We can also use `%in%` to filter with **dplyr**. You can therefore see the data from New York and Texas like this: 

```{r}
head(filter(murders, state %in% c("New York", "Texas")))
```

Create a new data frame called `murders_nw` with only the states from the Northeast and the West. How many states are in this category? 

```{r}
murders_nw <- filter(murders, region %in% c("Northeast", "West"))
nrow(murders_nw)
```



7. Suppose you want to live in the Northeast or West **and** want the murder rate to be less than 1. We want to see the data for the states satisfying these options. Note that you can use logical operators with filter. Here is an example in which we filter to keep only small states in the Northeast region.

```{r}
filter(murders, population < 5000000 & region == "Northeast")
```

Make sure `murders` has been defined with rate and rank and still has all states. Create a table called `my_states` that contains rows for states satisfying both the conditions: it is in the Northeast or West and the murder rate is less than 1. Use `select` to show only the state name, the rate and the rank.

```{r}
my_state <- filter(murders, region %in% c("Northeast", "West") & rate < 1)
select(my_state, state, rate, rank) %>% show()
```



***

#5.5 The pipe : `%>%`

With **dplyr** we can perform a series of operations, for example select and then `filter`, by sending the results of one function to another using what is called the pipe operator: `%>%`. Some details are included below.

We wrote code above to show three variables (state, region, rate) for states that have murder rates below 0.71. To do this, we defined the intermediate object `new_table`. In **dplyr** we can write code that looks more like a description of what we want to do without intermediate objects:

<center>$original -> select -> filter$</center>

For such an operation, we can use the pip `%>%`. The code looks like this:

```{r}
murders %>% select(state, region, rate) %>% filter(rate <= 0.71)
```

This line of code is equivalent to the two lines of code above. What is going on here?

In general, the pipe sends the result of the left side of the pipe to be the first argument of the function on the right side of the pipe. Here is a very simple example:

```{r}
16 %>% sqrt()
```

We can continue to pipe values along:

```{r}
16 %>% sqrt() %>% log2()
```

The above statement is equivalent to `log2(sqrt(16))`.

Remember that the pipe sends values to the **first argument**, so we can define other arguments as if the first argument is already defined:

```{r}
16 %>% sqrt() %>% log(base=2)
```

Therefore, when using the pipe with data frames and **dplyr**, we no longer need to specify the required first argument since the **dplyr** functions we have described all take the data as the first argument. In the code we wrote:

```{r}
# first argument는 이미 선언되었다! 
murders %>% select(state, region, rate) %>% filter(rate<=0.71)
```

`murders` is the first argument of the `select` function, and the new data frame (formerly `new_table`) is the first argument of the `filter` function.

Note that the pipe works well with functions where the first argument is the input data. Functions in **tidyverse** packages like **dplyr** have this format and can be used easily with the pipe.



***

# 5.6 Exercise 

1. The pipe `%>%` can be used to perform operations sequentially without having to define intermediate objects. Start by redefining murder to include rate and rank.

```{r}
data(murders)
murders <- mutate(murders, rate = total / population * 100000, rank = rank(-rate))
```

In the solution to the previous exercise, we did the following:

```{r}
my_states <- filter(murders, region %in% c("Northeast", "West") & rate < 1)
select(my_states, state, rate, rank)
```

The pipe `%>%` permits us to perform both operations sequentially without having to define an intermediate variable `my_states`. We therefore could have mutated and selected in the same line like this:

```{r}
mutate(murders, rate =  total / population * 100000, rank = rank(-rate)) %>%
  select(state, rate, rank)
```

Notice that `select` no longer has a data frame as the first argument. The first argument is assumed to be the result of the operation conducted right before the `%>%`.

Repeat the previous exercise, but now instead of creating a new object, show the result and only include the state, rate, and rank columns. Use a pipe `%>%` to do this in just one line.

```{r}
murders %>% filter(region %in% c("Northwest", "West") & rate < 1) %>% select(state, rate, rank) 
```



2. Reset `murders` to the original table by using `data(murders)`. Use a pipe to create a new data frame called `my_states` that considers only states in the Northeast or West which have a murder rate lower than 1, and contains only the state, rate and rank columns. The pipe should also have four components separated by three `%>%`. The code should look something like this:

```{r}
#pipe로 연결된 변수를 다른 변수에 바로 저장할 수 있다.
#흥미로운 점 : 아래의 mutate에서, rate 변수가 완전히 생성되기 전에 rank 함수에서 rate를 받을 수 있다. 아무래도 함수 내에서 argument의 순서가 영향을 미치는 듯. 

data(murders)
my_states <- murders %>% mutate(rate = total / population * 100000, rank = rank(-rate)) %>% filter(region %in% c("Northeast", "West"), rate < 1) %>% select(state, rate, rank)
my_states
```